---
title: "COO - binomial LR from Young et al.2018"
author: "Marta Interlandi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup_environment, message=FALSE, warning=FALSE}
library(gelnet)
library(dplyr)
library(biomaRt)
library(Seurat)
library(pheatmap)
library(Hmisc)
library(glmnet)
library(ggplot2)
library(cowplot)
library(foreach)
library(doMC)
library(WriteXLS)
library(plyr)


setwd("/marta_home/scRNAseq_Ranalysis/Seurat_v3/COO-logisticReg/binomialLR")

load("/marta_home/scRNAseq_Ranalysis/Seurat_v3/COO-logisticReg/binomialLR/workspace.RData")

```

# Collect genes to exclude from the model construction

```{r vrnbigobn}

## Exclude specific genes from the model construction
ensembl <- useMart("ensembl", dataset="mmusculus_gene_ensembl")

# Get mouse genes associated to mitochondrion and child terms
mito.genes.go <- getBM(attributes=c('mgi_symbol'),
                   filters = 'go', values = 'GO:0005739', mart = ensembl)

# Ribosome 
ribo.genes.go <- getBM(attributes=c('mgi_symbol'),
                   filters = 'go', values = 'GO:0005840', mart = ensembl)

# Cell cycle 
cellcycle.genes.go <- getBM(attributes=c('mgi_symbol'),
                   filters = 'go', values = 'GO:0007049', mart = ensembl)

go.genes2exclude <- unique(c(mito.genes.go$mgi_symbol, ribo.genes.go$mgi_symbol, cellcycle.genes.go$mgi_symbol))




```

# Read embryo reference dataset

```{r vfopbde}
embryo_ref <- readRDS(file = "/marta_home/scRNAseq_Ranalysis/Seurat_v3/Embryos/published_data/v3/embryo_ref_int.rds")
DefaultAssay(embryo_ref) <- "RNA"

celltypes.short <- as.character(embryo_ref$celltype)
celltypes.short <- plyr::mapvalues(celltypes.short, from = c("Blood progenitors 1", "Blood progenitors 2", "Erythroid1", "Erythroid2", "Erythroid3", "Endothelial cells", "Intermediate Mesoderm", "Notochord cells", "Notochord", "Cardiac muscle lineages"), to = c("Blood progenitors", "Blood progenitors", "Erythroid", "Erythroid", "Erythroid", "Endothelium", "Intermediate mesoderm", "Notochord late", "Notochord early", "Cardiomyocytes"))
embryo_ref$celltype.short <- factor(celltypes.short)
celltypes <- sort(unique(embryo_ref$celltype.short))

table(embryo_ref$celltype.short)
# Excluding lens
embryo_ref <- subset(embryo_ref, subset = celltype.short != "Lens")

embryo_ref$celltype.short <- factor(as.character(embryo_ref$celltype.short))

table(embryo_ref$celltype.short)

embryo <- embryo_ref[!(rownames(embryo_ref) %in% go.genes2exclude),]
embryo_mat <- GetAssayData(embryo, slot = "data", assay = "RNA")
dim(embryo_mat)
# keeping genes that are expressed in at least 100 cells
e <- CreateSeuratObject(embryo_mat, min.cells = 100)

embryo_mat <- GetAssayData(e, slot = "data", assay = "RNA")
dim(embryo_mat)
```



## Logistic regression adapted from Young et al.2018 https://github.com/constantAmateur/scKidneyTumors

```{r vmfiobgdhfd}
getPopulationOffset = function(y){
  if(!is.factor(y))
    y=factor(y)
  if(length(levels(y))!=2)
    stop("y must be a two-level factor")
  off = sum(y==levels(y)[2])/length(y)
  off = log(off/(1-off))
  return(rep(off,length(y)))
}


#' Do the OvR fit for every variable.  This just does a simple CV (cross-validation) selection of regularisation amount.  Far from ideal, but should be good enough for the main conclusions.
multinomialFitCV = function(x,y,nParallel=1,...){
  fits = list()
  if(nParallel>1)
    registerDoMC(cores=nParallel)
  #Do them in order of size
  marks = names(sort(table(as.character(y))))
  for(mark in marks){
    message(sprintf("Fitting model for variable %s",mark))
    fac = factor(y==mark)
    #The two main modes of failure are too few positives and errors constructing lambda.  These should be handled semi-gracefully
    fits[[mark]] = tryCatch(
      cv.glmnet(x,fac,offset=getPopulationOffset(fac),family='binomial',intercept=FALSE,alpha=0.8,nfolds=10,type.measure='class',parallel=nParallel>1,...),
      error = function(e) {
        tryCatch(
          cv.glmnet(x,fac,offset=getPopulationOffset(fac),family='binomial',intercept=FALSE,alpha=0.8,nfolds=10,type.measure='class',parallel=nParallel>1,lambda=exp(seq(-10,-3,length.out=100)),...),
          error = function(e) {
            warning(sprintf("Could not fit model for variable %s",mark))
            return(NULL)
          })
      })
  }
  return(fits)
}


# Training data (embryo cells)
trainDat <- t(embryo_mat)
dim(trainDat)

classes <- as.factor(embryo$celltype.short)


fitEmbryo <- multinomialFitCV(trainDat, classes, nParallel=40)

saveRDS(fitEmbryo, file = "lr_embryo_0602.rds")


```


## Prediction
##### SHH


```{r exc}
# get genes that are part of the glm model
glm.genes <- rownames(fitEmbryo[[1]]$glmnet.fit$beta)

# Tumor: SHH IC
tum_shh <- readRDS(file = "/marta_home/scRNAseq_Ranalysis/Seurat_v3/mouseATRT/v3/ATRT_IC_SHH/shh_tumor_cells.rds")

tum_shh_mat <- GetAssayData(tum_shh, slot = "data", assay = "RNA")
dim(tum_shh_mat)

add_missing_genes <- function(tum_mat, glm.genes){
  # add missing genes to tumor matrix (with 0 expression)
  # return test data
  genes.missing <- glm.genes[!(glm.genes %in% rownames(tum_mat))]
  zero_mat <- matrix(0, nrow = length(genes.missing), ncol = ncol(tum_mat))
  rownames(zero_mat) <- genes.missing
  colnames(zero_mat) <- colnames(tum_mat)
  tum_shh_mat_expanded <- rbind(tum_mat, zero_mat)
  testDat <- t(tum_shh_mat_expanded[glm.genes,]) 
  return(testDat)
}

testDat <- add_missing_genes(tum_shh_mat, glm.genes)
dim(testDat)

#Predict on tumor data
preds.shh = list()
for(ct in levels(classes)){
  message(sprintf("Predicting probabilities for cluster %s", ct))
  preds.shh[[ct]] = predict(fitEmbryo[[ct]], newx = testDat, s = 'lambda.1se', newoffset = rep(0,nrow(testDat)), 
                              type = "response")
}

# preds.shh.se = list()
# for(ct in levels(classes)){
#   message(sprintf("Predicting probabilities for cluster %s", ct))
#   preds.shh.se[[ct]] = predict(fitEmbryo[[ct]], newx = testDat, s = 'lambda.1se', newoffset = rep(0,nrow(testDat)), 
#                               type = "response", se.fit = TRUE)
# }

# Calculate mean probability for each cluster
prob_mean <- as.matrix(sort(unlist(lapply(preds.shh, mean))))
rownames(prob_mean)

pheatmap(prob_mean, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, filename = "SHH_IC_LogReg_probability.tiff")

prob.mat <- matrix(unlist(preds.shh), ncol = length(levels(classes)), byrow = F)
colnames(prob.mat) <- levels(classes)
rownames(prob.mat) <- colnames(tum_shh)
annotation <- data.frame(cluster = tum_shh$integrated_snn_res.0.5, row.names = colnames(tum_shh))

pheatmap(t(prob.mat), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "SHH_IC_LogReg_probability_allcells.tiff")

saveRDS(preds.shh, file = "SHH_preds.rds")

## Remove ct for which the model fails
prob_mean_nofail <- as.matrix(prob_mean[!(rownames(prob_mean) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early")),])
pheatmap(prob_mean_nofail, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, show_rownames = T, filename = "SHH_IC_LogReg_probability_nofail.tiff")

prob.mat.nofail <- prob.mat[, !(colnames(prob.mat) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early"))]
pheatmap(t(prob.mat.nofail), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "SHH_IC_LogReg_probability_allcells_nofail.tiff")

##Heatmap all cells ordered by cluster
rownames(prob.mat.nofail)
clusters <- as.character(tum_shh$integrated_snn_res.0.5)
names(clusters) <- colnames(tum_shh)
clusters <- clusters[order(clusters)]

prob.mat.nofail.ord <- t(prob.mat.nofail[names(clusters),])
# order celltypes by avg probability ascending order
prob.mat.nofail.ord <- prob.mat.nofail.ord[rownames(prob_mean_nofail),]

pheatmap(prob.mat.nofail.ord, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "SHH_IC_LogReg_probability_allcells_nofailBYCLUST.tiff")

## aggregate cluster by computing average
heat_mat <- prob.mat.nofail.ord
mat <- as.data.frame(t(heat_mat))

mat$cluster <- clusters

mean.mat <- aggregate(. ~ cluster, mat, mean)
rownames(mean.mat) <- mean.mat$cluster

annotation_col <- data.frame(annotation= mean.mat$cluster, row.names = rownames(mean.mat))


mean.mat <- mean.mat[,-1]
mean.mat <- t(as.matrix(mean.mat))

pheatmap(mean.mat, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation_col, filename = "SHH_IC_LogReg_probability_allcells_nofailBYCLUSTmean.tiff")


# Coefficients for mid/hindbrain
midhind.coef <- as.matrix(coef(fitEmbryo[["Midbrain/Hindbrain"]]))
dim(midhind.coef)
midhind.predictors <- rownames(midhind.coef)[midhind.coef != 0]
midhind.coef.df <- data.frame(predictors = midhind.predictors, coefficient = midhind.coef[midhind.coef != 0])
WriteXLS(midhind.coef.df, ExcelFileName = "MidHindbrain_predictors.xlsx", row.names = F)

saveRDS(midhind.predictors, "midhind_predictors_genes.rds")


```

##### MYC IC


```{r exc}
# get genes that are part of the glm model
glm.genes <- rownames(fitEmbryo[[1]]$glmnet.fit$beta)

tum_myc_ic <- readRDS(file = "/marta_home/scRNAseq_Ranalysis/Seurat_v3/mouseATRT/v3/ATRT_IC_MYC/myc_ic_tumor_cells.rds")

myc_ic_mat <- GetAssayData(tum_myc_ic, slot = "data", assay = "RNA")
dim(myc_ic_mat)
testDat <- add_missing_genes(myc_ic_mat, glm.genes)

#Predict on tumor data
preds.myc.ic = list()
for(ct in levels(classes)){
  message(sprintf("Predicting probabilities for cluster %s", ct))
  preds.myc.ic[[ct]] = predict(fitEmbryo[[ct]], newx = testDat, s = 'lambda.1se', newoffset = rep(0,nrow(testDat)), 
                              type = "response")
}



prob_mean <- as.matrix(sort(unlist(lapply(preds.myc.ic, mean))))

pheatmap(prob_mean, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, filename = "MYC_IC_LogReg_probability.tiff")

prob.mat <- matrix(unlist(preds.myc.ic), ncol = length(levels(classes)), byrow = F)
colnames(prob.mat) <- levels(classes)
rownames(prob.mat) <- colnames(tum_myc_ic)
annotation <- data.frame(cluster = tum_myc_ic$new_clustering, row.names = colnames(tum_myc_ic))

pheatmap(t(prob.mat), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_IC_LogReg_probability_allcells.tiff")


saveRDS(preds.myc.ic, file = "MYC_IC_preds.rds")

## Remove ct for which the model fails
prob_mean_nofail <- as.matrix(prob_mean[!(rownames(prob_mean) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early")),])
pheatmap(prob_mean_nofail, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, show_rownames = T, filename = "MYC_IC_LogReg_probability_nofail.tiff")

prob.mat.nofail <- prob.mat[, !(colnames(prob.mat) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early"))]
pheatmap(t(prob.mat.nofail), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_IC_LogReg_probability_allcells_nofail.tiff")

##Heatmap all cells ordered by cluster
rownames(prob.mat.nofail)
clusters <- as.character(tum_myc_ic$new_clustering)
names(clusters) <- colnames(tum_myc_ic)
clusters <- clusters[order(clusters)]

prob.mat.nofail.ord <- t(prob.mat.nofail[names(clusters),])
# order celltypes by avg probability ascending order
prob.mat.nofail.ord <- prob.mat.nofail.ord[rownames(prob_mean_nofail),]

pheatmap(prob.mat.nofail.ord, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_IC_LogReg_probability_allcells_nofailBYCLUST.tiff")

## aggregate cluster by computing average
heat_mat <- prob.mat.nofail.ord
mat <- as.data.frame(t(heat_mat))

mat$cluster <- clusters

mean.mat <- aggregate(. ~ cluster, mat, mean)
rownames(mean.mat) <- mean.mat$cluster

annotation_col <- data.frame(annotation= mean.mat$cluster, row.names = rownames(mean.mat))


mean.mat <- mean.mat[,-1]
mean.mat <- t(as.matrix(mean.mat))

pheatmap(mean.mat, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation_col, filename = "MYC_IC_LogReg_probability_allcells_nofailBYCLUSTmean.tiff")



```

##### MYC EC


```{r exc}
# get genes that are part of the glm model
glm.genes <- rownames(fitEmbryo[[1]]$glmnet.fit$beta)

tum_myc_ec <- readRDS(file = "/marta_home/scRNAseq_Ranalysis/Seurat_v3/mouseATRT/v3/MRT_EC_MYC/myc_ec_tumor_cells.rds")
tum_myc_ec_mat <- GetAssayData(tum_myc_ec, slot = "data", assay = "RNA")

testDat <- add_missing_genes(tum_myc_ec_mat, glm.genes)
dim(testDat)

#Predict on tumour data
preds.myc.ec = list()
for(mark in levels(classes)){
  message(sprintf("Predicting probabilities for cluster %s", mark))
  preds.myc.ec[[mark]] = predict(fitEmbryo[[mark]], newx = testDat, s = 'lambda.1se', newoffset = rep(0,nrow(testDat)), 
                              type = "response")
}


prob_mean <- as.matrix(sort(unlist(lapply(preds.myc.ec, mean))))

pheatmap(prob_mean, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, filename = "MYC_EC_LogReg_probability.tiff")

prob.mat <- matrix(unlist(preds.myc.ec), ncol = length(levels(classes)), byrow = F)
colnames(prob.mat) <- levels(classes)
rownames(prob.mat) <- colnames(tum_myc_ec)

annotation <- data.frame(cluster = tum_myc_ec$new_clustering, row.names = colnames(tum_myc_ec))

pheatmap(t(prob.mat), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation,  filename = "MYC_EC_LogReg_probability_allcells.tiff")


saveRDS(preds.myc.ec, file = "MYC_EC_preds.rds")

## Remove ct for which the model fails
prob_mean_nofail <- as.matrix(prob_mean[!(rownames(prob_mean) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early")),])
pheatmap(prob_mean_nofail, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, show_rownames = T, filename = "MYC_EC_LogReg_probability_nofail.tiff")

prob.mat.nofail <- prob.mat[, !(colnames(prob.mat) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early"))]
pheatmap(t(prob.mat.nofail), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_EC_LogReg_probability_allcells_nofail.tiff")



##Heatmap all cells ordered by cluster
rownames(prob.mat.nofail)
clusters <- as.character(tum_myc_ec$new_clustering)
names(clusters) <- colnames(tum_myc_ec)
clusters <- clusters[order(clusters)]

prob.mat.nofail.ord <- t(prob.mat.nofail[names(clusters),])
# order celltypes by avg probability ascending order
prob.mat.nofail.ord <- prob.mat.nofail.ord[rownames(prob_mean_nofail),]

pheatmap(prob.mat.nofail.ord, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_EC_LogReg_probability_allcells_nofailBYCLUST.tiff")


# Adding info on Sox2/Rosa-cre models
model <- as.character(tum_myc_ec$sample)
table(model)
model <- mapvalues(model, from = c("ros1148", "sox568", "sox593"), to = c("Rosa26-cre", "Sox2-cre", "Sox2-cre"))
names(model) <- colnames(tum_myc_ec)

annotation_model <- annotation
annotation_model$model <- model

pheatmap(prob.mat.nofail.ord, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation_model, filename = "MYC_EC_LogReg_probability_allcells_nofailBYCLUST_model.tiff")


## aggregate cluster by computing average
heat_mat <- prob.mat.nofail.ord
mat <- as.data.frame(t(heat_mat))

mat$cluster <- clusters

mean.mat <- aggregate(. ~ cluster, mat, mean)
rownames(mean.mat) <- mean.mat$cluster

annotation_col <- data.frame(annotation= mean.mat$cluster, row.names = rownames(mean.mat))


mean.mat <- mean.mat[,-1]
mean.mat <- t(as.matrix(mean.mat))

pheatmap(mean.mat, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation_col, filename = "MYC_EC_LogReg_probability_allcells_nofailBYCLUSTmean.tiff")



# Coefficients for Connective tissue progenitors
coef <- as.matrix(coef(fitEmbryo[["Connective tissue progenitors"]]))
predictors <- rownames(coef)[coef != 0]
coef.df <- data.frame(predictors = predictors, coefficient = coef[coef != 0])
WriteXLS(coef.df, ExcelFileName = "Connective_tis_prog_predictors.xlsx", row.names = F)

saveRDS(predictors, "Connective_tis_prog_predictors_genes.rds")



```


##### Spinal


```{r exc}
# get genes that are part of the glm model
glm.genes <- rownames(fitEmbryo[[1]]$glmnet.fit$beta)

tum_spinal <- readRDS(file = "/marta_home/scRNAseq_Ranalysis/Seurat_v3/mouseATRT/v3/ATRT_spinal_MYC/myc_spinal_tumor_cells.rds")
tum_spinal_mat <- GetAssayData(tum_spinal, slot = "data", assay = "RNA")

testDat <- add_missing_genes(tum_spinal_mat, glm.genes)
dim(testDat)

#Predict on tumour data
preds.myc.sp = list()
for(mark in levels(classes)){
  message(sprintf("Predicting probabilities for cluster %s", mark))
  preds.myc.sp[[mark]] = predict(fitEmbryo[[mark]], newx = testDat, s = 'lambda.1se', newoffset = rep(0,nrow(testDat)), 
                              type = "response")
}



prob_mean <- as.matrix(sort(unlist(lapply(preds.myc.sp, mean))))

pheatmap(prob_mean, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, filename = "MYC_SPIN_LogReg_probability.tiff")

prob.mat <- matrix(unlist(preds.myc.sp), ncol = length(levels(classes)), byrow = F)
colnames(prob.mat) <- levels(classes)
rownames(prob.mat) <- colnames(tum_spinal)

annotation <- data.frame(cluster = tum_spinal$new_clustering, row.names = colnames(tum_spinal))

pheatmap(t(prob.mat), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation,  filename = "MYC_SPIN_LogReg_probability_allcells.tiff")


saveRDS(preds.myc.sp, file = "MYC_SPIN_preds.rds")

## Remove ct for which the model fails
prob_mean_nofail <- as.matrix(prob_mean[!(rownames(prob_mean) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early")),])
pheatmap(prob_mean_nofail, cluster_rows = F, cluster_cols = F, cellwidth = 10, cellheight = 10, show_rownames = T, filename = "MYC_SPIN_LogReg_probability_nofail.tiff")

prob.mat.nofail <- prob.mat[, !(colnames(prob.mat) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early"))]
pheatmap(t(prob.mat.nofail), cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_SPIN_LogReg_probability_allcells_nofail.tiff")

##Heatmap all cells ordered by cluster
rownames(prob.mat.nofail)
clusters <- as.character(tum_spinal$new_clustering)
names(clusters) <- colnames(tum_spinal)
clusters <- clusters[order(clusters)]

prob.mat.nofail.ord <- t(prob.mat.nofail[names(clusters),])
# order celltypes by avg probability ascending order
prob.mat.nofail.ord <- prob.mat.nofail.ord[rownames(prob_mean_nofail),]

pheatmap(prob.mat.nofail.ord, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation, filename = "MYC_SPIN_LogReg_probability_allcells_nofailBYCLUST.tiff")

## aggregate cluster by computing average
heat_mat <- prob.mat.nofail.ord
mat <- as.data.frame(t(heat_mat))

mat$cluster <- clusters

mean.mat <- aggregate(. ~ cluster, mat, mean)
rownames(mean.mat) <- mean.mat$cluster

annotation_col <- data.frame(annotation= mean.mat$cluster, row.names = rownames(mean.mat))


mean.mat <- mean.mat[,-1]
mean.mat <- t(as.matrix(mean.mat))

pheatmap(mean.mat, cluster_rows = F, cluster_cols = F, show_rownames = T, show_colnames = F, annotation_col = annotation_col, filename = "MYC_SPIN_LogReg_probability_allcells_nofailBYCLUSTmean.tiff")


# Coefficients for ExE ectoderm
coef <- as.matrix(coef(fitEmbryo[["ExE ectoderm"]]))
predictors <- rownames(coef)[coef != 0]
coef.df <- data.frame(predictors = predictors, coefficient = coef[coef != 0])
WriteXLS(coef.df, ExcelFileName = "ExE_ectoderm_predictors.xlsx", row.names = F)

saveRDS(predictors, "ExE_ectoderm_predictors_genes.rds")
```




## Permutation test on predicted scores

We randomly permute (1,000 times) the scores per cell across cell types and calculate the average score per celltype each time. Then we compare it to the actual mean score. The exact pvalue (one-sided) would be the proportion of the means that are as or higher than the actual mean.

```{r ruiht}
# Actual means

getactualMean <- function(predictions){
  prob_mean <- as.matrix(sort(unlist(lapply(predictions, mean))))

  ## Remove ct for which the model fails
  prob_mean_nofail <- as.matrix(prob_mean[!(rownames(prob_mean) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early")),])
  return(prob_mean_nofail)
}

shh.actualMeans <- getactualMean(preds.shh)
myc.ic.actualMeans <- getactualMean(preds.myc.ic)
myc.ec.actualMeans <- getactualMean(preds.myc.ec)
myc.sp.actualMeans <- getactualMean(preds.myc.sp)


getProbMat <- function(predictions, classes){
  prob.mat <- matrix(unlist(predictions), ncol = length(levels(classes)), byrow = F)
  colnames(prob.mat) <- levels(classes)

  ## Remove ct for which the model fails
  prob.mat.nofail <- prob.mat[, !(colnames(prob.mat) %in% c("Melanocytes","White blood cells", "Megakaryocytes", "Notochord early"))]
  return(prob.mat.nofail)
}

# cells by rows, celltypes by cols
shh.ProbMat <- getProbMat(preds.shh, classes)
myc.ic.ProbMat <- getProbMat(preds.myc.ic, classes)
myc.ec.ProbMat <- getProbMat(preds.myc.ec, classes)
myc.sp.ProbMat <- getProbMat(preds.myc.sp, classes)

getPvalue_df <- function(probMat, actualMeans){

    permMat_mean <- list()
    set.seed(123)
    for(np in seq_len(999)){
        # shuffle cols of original matrix (celltypes)
        shufMat <- probMat[,sample(colnames(probMat), ncol(probMat), 
                                   replace = FALSE)]
        colnames(shufMat) <- colnames(probMat)
        permMat_mean[[np]] <- colMeans(shufMat)
    }
    
    # calculate empirical pvalue
    emp_pvalue <- numeric(nrow(actualMeans))
    for(ct in seq_len(nrow(actualMeans))){
        permMeans <- unlist(lapply(permMat_mean, function(x) x[ct]))
        emp_pvalue[ct] <- (1 + sum(permMeans >= actualMeans[ct]))/1000
    }
    
    pvalue_df <- data.frame(celltype = rownames(actualMeans), p_val = emp_pvalue)
    pvalue_df <- pvalue_df %>%
      arrange(p_val)
    return(pvalue_df)
}
    
shh.pval.df <- getPvalue_df(shh.ProbMat, shh.actualMeans)
myc.ic.pval.df <- getPvalue_df(myc.ic.ProbMat, myc.ic.actualMeans)
myc.ec.pval.df <- getPvalue_df(myc.ec.ProbMat, myc.ec.actualMeans)
myc.sp.pval.df <- getPvalue_df(myc.sp.ProbMat, myc.sp.actualMeans)

WriteXLS::WriteXLS(list(SHH = shh.pval.df, MYC_IC = myc.ic.pval.df, eRT = myc.ec.pval.df, MYC_SPINAL = myc.sp.pval.df), ExcelFileName = "permutationtest_pvalues_subgroups.xlsx")
 



```

